---
title: "AT1_A_MODELLING_1"
author: "Joshua McCarthy"
date: "16/04/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, eval = FALSE}
install.packages("lubridate")
install.packages("MuMIn")
install.packages("glmnet")
```


```{r}
library(lubridate)
library(tidyverse)
library(MuMIn)
library(glmnet)
```



```{r Prep}
#import
transactions <- read_csv("transactions.csv")

# convert dates to dates
transactions$date <- as.Date(transactions$date, format = "%d/%m/%y")

#change variable name to avoid conflicts
tname <- names(transactions)
tname[1] <- "trdate"
names(transactions) <- tname

# convert neccessary variables to factors
transactions$customer_id <- as.factor(transactions$customer_id) 
transactions$industry <- as.factor(transactions$industry)
transactions$location <- as.factor(transactions$location)
```

# 2. Basic Model Fitting

## a. Creating the model

### i. Create an aggregated data set using the fields date, industry and location, with a mean of monthly_amount.

```{r}
aggregated_transactions <- group_by(transactions, trdate, location, industry) %>%
  summarise(mean_amount = mean(monthly_amount))
```

### ii. Create a line plot of the variable monthly_amount for industry = 1 and location = 1.  Note the seasonality by month in this time series. 

```{r}
aggregated_transactions$trmonth <- lubridate::month(aggregated_transactions$trdate, label = TRUE, abbr = TRUE)
```

```{r}
# Output date list
trdates_out <- data.frame(unique(aggregated_transactions$trdate))
names(trdates_out) <- "trdate"
# trdates_out$trdate <- format(as.Date(trdates_out$trdate), "%Y-%m-%d")
write_csv(trdates_out, "trdates_out.csv")
```

```{r}
a_i1_l1 <- aggregated_transactions %>%
  filter(location == 1, industry == 1)

a_i1_l1_plot <- ggplot(a_i1_l1 , aes( x = trdate, y = mean_amount)) + geom_line() + geom_point() + geom_text(aes(label = trmonth, vjust=-1))

a_i1_l1_plot
```

it looks like the usual seasonailiy is offset in 2016, why?
 A. Australian double dissolution federal election?

### iii. For industry= 1 and location = 1, train a linear regression model with monthly_amount as the target. 

Intial Model test
```{r}
a_i1_l1_lm_bad <- lm(mean_amount ~ trdate, a_i1_l1)
summary(a_i1_l1_lm_bad)
plot(a_i1_l1_lm_bad)

a11lm_predict <- data.frame(mean_amount_pred = predict(a_i1_l1_lm_bad, a_i1_l1), trdate = a_i1_l1$trdate)

a_i1_l1_plot + geom_line(color ="red", data = a11lm_predict, aes(y = mean_amount_pred))
``` 

#### i. Note 1 :Remember that time is very important in this model, so be sure to include variable(s) for the time sequence.

```{r}
# EDA
a_i1_l1_filt <- a_i1_l1 %>%
  data.frame() %>%
  select(mean_amount)
ts_a_i1_l1 <- ts(a_i1_l1_filt, start = decimal_date(as.Date("2013-01-01")), frequency = 12)
plot(ts_a_i1_l1)
components_a_i1_l1 <- decompose(ts_a_i1_l1)
plot(components_a_i1_l1)
```

```{r}
a_i1_l1 <- a_i1_l1 %>%
  data.frame() %>%
  select(-location, -industry)
```


```{r}
a_i1_l1_lm_bad <- lm(mean_amount ~., a_i1_l1)
summary(a_i1_l1_lm_bad)
plot(a_i1_l1_lm_bad)

a11lm_predict <- data.frame(mean_amount_pred = predict(a_i1_l1_lm_bad, a_i1_l1), trdate = a_i1_l1$trdate)

a_i1_l1_plot + geom_line(color ="red", data = a11lm_predict, aes(y = mean_amount_pred))
```


```{r}
# calendar quarters
a_i1_l1$calquarters <- "cQ4"
  a_i1_l1[a_i1_l1$trmonth == "Jan" | a_i1_l1$trmonth == "Feb" | a_i1_l1$trmonth == "Mar" , "calquarters"] <- "cQ1"
  a_i1_l1[a_i1_l1$trmonth == "Apr" | a_i1_l1$trmonth == "May" | a_i1_l1$trmonth == "Jun" , "calquarters"] <- "cQ2"
  a_i1_l1[a_i1_l1$trmonth == "Jul" | a_i1_l1$trmonth == "Aug" | a_i1_l1$trmonth == "Sep" , "calquarters"] <- "cQ3"
a_i1_l1$calquarters <- as.ordered(a_i1_l1$calquarters)  
```

```{r}
# aus fiscal quarters
a_i1_l1$fisquarters <- "fQ4"
  a_i1_l1[a_i1_l1$trmonth == "Jul" | a_i1_l1$trmonth == "Aug" | a_i1_l1$trmonth == "Sep" , "fisquarters"] <- "fQ1"
  a_i1_l1[a_i1_l1$trmonth == "Oct" | a_i1_l1$trmonth == "Nov" | a_i1_l1$trmonth == "Dec" , "fisquarters"] <- "fQ2"
  a_i1_l1[a_i1_l1$trmonth == "Jan" | a_i1_l1$trmonth == "Feb" | a_i1_l1$trmonth == "Mar" , "fisquarters"] <- "fQ3"
a_i1_l1$fisquarters <- as.ordered(a_i1_l1$fisquarters)  
```

```{r}
# Seasons
a_i1_l1$seasons <- "spring"
  a_i1_l1[a_i1_l1$trmonth == "Dec" | a_i1_l1$trmonth == "Jan" | a_i1_l1$trmonth == "Feb" , "seasons"] <- "summer"
  a_i1_l1[a_i1_l1$trmonth == "Mar" | a_i1_l1$trmonth == "Apr" | a_i1_l1$trmonth == "May" , "seasons"] <- "autumn"
  a_i1_l1[a_i1_l1$trmonth == "Jun" | a_i1_l1$trmonth == "Jul" | a_i1_l1$trmonth == "Aug" , "seasons"] <- "winter"
a_i1_l1$seasons <- as.ordered(a_i1_l1$seasons)
```

```{r}
a_i1_l1_lm_bad <- lm(mean_amount ~ trdate + calquarters + seasons + fisquarters, a_i1_l1)
summary(a_i1_l1_lm_bad)
plot(a_i1_l1_lm_bad)

a11lm_predict <- data.frame(mean_amount_pred = predict(a_i1_l1_lm_bad, a_i1_l1), trdate = a_i1_l1$trdate)

a_i1_l1_plot + geom_line(color ="red", data = a11lm_predict, aes(y = mean_amount_pred))
```
```{r}
a_i1_l1_lm_bad <- lm(mean_amount ~ trdate + fisquarters + trmonth, a_i1_l1)
summary(a_i1_l1_lm_bad)
plot(a_i1_l1_lm_bad)

a11lm_predict <- data.frame(mean_amount_pred = predict(a_i1_l1_lm_bad, a_i1_l1), trdate = a_i1_l1$trdate)

a_i1_l1_plot + geom_line(color ="red", data = a11lm_predict, aes(y = mean_amount_pred))
```

```{r}
# inflation 
# https://www.statista.com/statistics/271845/inflation-rate-in-australia/
inflation <- tibble(year = c(2013, 2014, 2015, 2016), inflperc = c(2.45, 2.51, 1.51, 1.25))

# gdp growth
# https://data.worldbank.org/indicator/NY.GDP.MKTP.KD.ZG?end=2016&locations=AU&start=2013
gdpgrowth <- tibble(year = c(2013, 2014, 2015, 2016), growthperc = c(2.585,2.533,2.193,2.771))
gdpgrowth <- gdpgrowth %>%
  slice(rep(1:n(), each = 12))

gdpgrowth$growthperc <- gdpgrowth$growthperc/12/100
gdpgrowth <- gdpgrowth[-48,]
gdpgrowth$growthperccumul <- gdpgrowth$growthperc
gdpgrowth <- data.frame(gdpgrowth)

for (inter in 2:nrow(gdpgrowth)) {
  gdpgrowth[inter,3] <- gdpgrowth[[inter,2]] + gdpgrowth[[inter-1,3]]
}
```

```{r}
a_i1_l1$gdp <- gdpgrowth$growthperccumul
plot(a_i1_l1$trdate, a_i1_l1$gdp)
```


```{r}
a_i1_l1_lm_bad <- lm(mean_amount ~ trmonth + gdp, a_i1_l1)
summary(a_i1_l1_lm_bad)
plot(a_i1_l1_lm_bad)

a11lm_predict <- data.frame(mean_amount_pred = predict(a_i1_l1_lm_bad, a_i1_l1), trdate = a_i1_l1$trdate)

a_i1_l1_plot + geom_line(color ="red", data = a11lm_predict, aes(y = mean_amount_pred))
```


```{r}
# CPI
# https://www.rba.gov.au/inflation/measures-cpi.html
df <- data.frame(cpi = c(102.0 ,102.4 ,102.8 ,104.0 ,104.8 ,105.4 ,105.9 ,106.4 ,106.6 ,106.8 ,107.5 ,108.0 ,108.4 , 108.2 ,108.6 ,109.4 ,110.0 ), d = (1:17))

plot(df$d, df$cpi)
```


```{r}
# points interpolated to match dataset in CPIData_editing
CPI_tdates <- read_csv("CPI_tdates.csv")

a_i1_l1$CPI <- CPI_tdates$trindex_predict 
a_i1_l1_lm_bad <- lm(mean_amount ~ CPI, a_i1_l1)
summary(a_i1_l1_lm_bad)
plot(a_i1_l1_lm_bad)

a11lm_predict <- data.frame(mean_amount_pred = predict(a_i1_l1_lm_bad, a_i1_l1), trdate = a_i1_l1$trdate)

a_i1_l1_plot + geom_line(color ="red", data = a11lm_predict, aes(y = mean_amount_pred))

```



```{r}
a_i1_l1_lm_bad <- lm(mean_amount ~ trdate + CPI + trmonth, a_i1_l1)
summary(a_i1_l1_lm_bad)
plot(a_i1_l1_lm_bad)

a11lm_predict <- data.frame(mean_amount_pred = predict(a_i1_l1_lm_bad, a_i1_l1), trdate = a_i1_l1$trdate)

a_i1_l1_plot + geom_line(color ="red", data = a11lm_predict, aes(y = mean_amount_pred))
```


```{r}
# vs USD

# vs EURO

# vs YUAN
```


```{r}
a_i1_l1_lm_bad <- lm(mean_amount ~ trmonth + gdp, a_i1_l1)
summary(a_i1_l1_lm_bad)
plot(a_i1_l1_lm_bad)

a11lm_predict <- data.frame(mean_amount_pred = predict(a_i1_l1_lm_bad, a_i1_l1), trdate = a_i1_l1$trdate)

a_i1_l1_plot + geom_line(color ="red", data = a11lm_predict, aes(y = mean_amount_pred))
```



* Holidays
* School Holidays
```{r}
# School Holidays
a_i1_l1$shols <- "0"
  a_i1_l1[a_i1_l1$trmonth == "Apr" , "shols"] <- 2.5
  a_i1_l1[a_i1_l1$trmonth == "Jul" , "shols"] <- 2
  a_i1_l1[a_i1_l1$trmonth == "Sep" , "shols"] <- 1
  a_i1_l1[a_i1_l1$trmonth == "Oct" , "shols"] <- 1
  a_i1_l1[a_i1_l1$trmonth == "Dec" , "shols"] <- 2.5
  a_i1_l1[a_i1_l1$trmonth == "Jan" , "shols"] <- 4
a_i1_l1$shols <- as.numeric(a_i1_l1$shols)
  
  summary(a_i1_l1$shols)
```


```{r}
a_i1_l1_lm_bad <- lm(mean_amount ~ trdate + shols + trmonth, a_i1_l1)
summary(a_i1_l1_lm_bad)
plot(a_i1_l1_lm_bad)

a11lm_predict <- data.frame(mean_amount_pred = predict(a_i1_l1_lm_bad, a_i1_l1), trdate = a_i1_l1$trdate)

a_i1_l1_plot + geom_line(color ="red", data = a11lm_predict, aes(y = mean_amount_pred))
```

```{r}
# Major hols; new year, easter, christmas

# Test as individual
```


* Index's?
** All ords
```{r}
ASX_All_Ordinaries <- read_csv("ASX All Ordinaries.csv")
ASX_All_Ordinaries_price <- ASX_All_Ordinaries %>%
  select(Date, Price)
a_i1_l1 <- merge(a_i1_l1, ASX_All_Ordinaries_price, by.x = "trdate", by.y = "Date", all.x = TRUE)
names(a_i1_l1)[names(a_i1_l1) == "Price"] <- "All_Ords"
```


** S&P ASX 20
```{r}
S_P_ASX_20 <- read_csv("S&P_ASX 20.csv")
S_P_ASX_20_price <- S_P_ASX_20 %>%
  select(Date, Price)
a_i1_l1 <- merge(a_i1_l1, S_P_ASX_20_price, by.x = "trdate", by.y = "Date", all.x = TRUE)
names(a_i1_l1)[names(a_i1_l1) == "Price"] <- "ASX_20"
```


** Small ords
```{r}
ASX_Small_Ordinaries <- read_csv("ASX Small Ordinaries.csv")
ASX_Small_Ordinaries_price <- ASX_Small_Ordinaries %>%
  select(Date, Price)
a_i1_l1 <- merge(a_i1_l1, ASX_Small_Ordinaries_price, by.x = "trdate", by.y = "Date", all.x = TRUE)
names(a_i1_l1)[names(a_i1_l1) == "Price"] <- "Small_Ords"
```

```{r}
a_i1_l1 %>%
  select(All_Ords, Small_Ords, ASX_20, mean_amount) %>%
  scale() %>%
  data.frame() %>% 
  cbind(a_i1_l1$trdate) %>%

ggplot(aes(x = a_i1_l1$trdate)) + 
  geom_line(color = "red", aes(y = All_Ords))  + 
  geom_line(color = "yellow", aes(y = Small_Ords)) +
  geom_line(color = "green", aes(y = ASX_20)) +
  geom_line(color = "black", aes(y = mean_amount))
```


```{r}
a_i1_l1_lm_bad <- lm(mean_amount ~ trdate + ASX_20 + trmonth, a_i1_l1)
summary(a_i1_l1_lm_bad)
plot(a_i1_l1_lm_bad)

a11lm_predict <- data.frame(mean_amount_pred = predict(a_i1_l1_lm_bad, a_i1_l1), trdate = a_i1_l1$trdate)

a_i1_l1_plot + geom_line(color ="red", data = a11lm_predict, aes(y = mean_amount_pred))
```


```{r}
a_i1_l1_lm_bad <- lm(mean_amount ~., a_i1_l1)
summary(a_i1_l1_lm_bad)
plot(a_i1_l1_lm_bad)

a11lm_predict <- data.frame(mean_amount_pred = predict(a_i1_l1_lm_bad, a_i1_l1), trdate = a_i1_l1$trdate)

a_i1_l1_plot + geom_line(color ="red", data = a11lm_predict, aes(y = mean_amount_pred))
```

recent election caused a change in the seasonality? how to account for it?

* Seasons
* Inflation
* Population growth
* Economic Growth
* Dollar

##### (Hint: on your plot you may see local trends like seasonality. Consider how you could craft a variable to capture this?).

##### You may also see a global upwards or downwards slope, could you craft a variable to capture this? Therefore there are two simple variables you could create to capture time.

##### Could you craft more complex ones, perhaps with polynomials to capture local or global trends? Experiment and see! 

```{r}
a_i1_l1_lm_bad <- lm(mean_amount ~ poly(trdate, degree = 2), data = a_i1_l1)
summary(a_i1_l1_lm_bad)
plot(a_i1_l1_lm_bad)

a11lm_predict <- data.frame(mean_amount_pred = predict(a_i1_l1_lm_bad, a_i1_l1), trdate = a_i1_l1$trdate)

a_i1_l1_plot + geom_line(color ="red", data = a11lm_predict, aes(y = mean_amount_pred))
```


```{r}
# test all variable combos
dredge_model <- lm(mean_amount ~., data = a_i1_l1,  na.action = "na.fail")
dout <- dredge(dredge_model, beta = "none", evaluate = TRUE)
# options(na.action = "na.omit")
```
```{r}
a_i1_l1_lm_bad <- lm(mean_amount ~ CPI + gdp + shols, data = a_i1_l1)
summary(a_i1_l1_lm_bad)
plot(a_i1_l1_lm_bad)

a11lm_predict <- data.frame(mean_amount_pred = predict(a_i1_l1_lm_bad, a_i1_l1), trdate = a_i1_l1$trdate)

a_i1_l1_plot + geom_line(color ="red", data = a11lm_predict, aes(y = mean_amount_pred))
```


```{r}
a_i1_l1_lm_bad <- lm(mean_amount ~ All_Ords + ASX_20 + CPI + gdp + Small_Ords + trdate + trmonth , data = a_i1_l1)
summary(a_i1_l1_lm_bad)
plot(a_i1_l1_lm_bad)

a11lm_predict <- data.frame(mean_amount_pred = predict(a_i1_l1_lm_bad, a_i1_l1), trdate = a_i1_l1$trdate)

a_i1_l1_plot + geom_line(color ="red", data = a11lm_predict, aes(y = mean_amount_pred))
```

```{r}
a_i1_l1_lm_bad <- lm(mean_amount ~ fisquarters + CPI + gdp + seasons+  shols , data = a_i1_l1)
summary(a_i1_l1_lm_bad)
plot(a_i1_l1_lm_bad)

a11lm_predict <- data.frame(mean_amount_pred = predict(a_i1_l1_lm_bad, a_i1_l1), trdate = a_i1_l1$trdate)

a_i1_l1_plot + geom_line(color ="red", data = a11lm_predict, aes(y = mean_amount_pred))
```

```{r}
a_i1_l1_lm_bad <- lm(mean_amount ~ poly(trdate, degree =2) * trmonth + ASX_20 , data = a_i1_l1)
summary(a_i1_l1_lm_bad)
plot(a_i1_l1_lm_bad)

a11lm_predict <- data.frame(mean_amount_pred = predict(a_i1_l1_lm_bad, a_i1_l1), trdate = a_i1_l1$trdate)

a_i1_l1_plot + geom_line(color ="red", data = a11lm_predict, aes(y = mean_amount_pred))
```
#### ii. Note 2: Carefully think about how you split your test and train sets. (Hint: Random is not appropriate!)

Initially thought training set should be the earlier years, however upon noticing skewing due to recent events I'll choose the latest 3 years and then test on the earlier

Splitting will be;
* Training set - 3 earliest years (01/01/2013 - 31/12/2015)
* Testing set - most recent year(01/01/2016 - end)

```{r}
a_i1_l1_trainset <- a_i1_l1 %>%
  filter(trdate <= "2015-12-31")

a_i1_l1_testset <- a_i1_l1 %>%
  filter(trdate > "2015-12-31")
```

```{r}
a_i1_l1_lm_train <- lm(mean_amount ~ poly(trdate, degree =2) + trmonth + ASX_20, data = a_i1_l1_trainset)
summary(a_i1_l1_lm_train)
plot(a_i1_l1_lm_train)

a11lm_train_predict <- data.frame(mean_amount_pred = predict(a_i1_l1_lm_train, a_i1_l1_trainset), trdate = a_i1_l1_trainset$trdate)

ggplot(a_i1_l1_trainset , aes( x = trdate, y = mean_amount)) + geom_line() + geom_point() + geom_text(aes(label = trmonth, vjust=-1)) + geom_line(color ="red", data = a11lm_train_predict, aes(y = mean_amount_pred))
```

```{r}
a_i1_l1_trainset <- a_i1_l1 %>%
  filter(trdate >= "2013-12-31")

a_i1_l1_testset <- a_i1_l1 %>%
  filter(trdate < "2013-12-31")
```

```{r}
a_i1_l1_lm_train <- lm(mean_amount ~ poly(trdate, degree =2) * shols + trmonth + CPI * gdp, data = a_i1_l1_trainset)
summary(a_i1_l1_lm_train)
plot(a_i1_l1_lm_train)

a11lm_train_predict <- data.frame(mean_amount_pred = predict(a_i1_l1_lm_train, a_i1_l1_trainset), trdate = a_i1_l1_trainset$trdate)

ggplot(a_i1_l1_trainset , aes( x = trdate, y = mean_amount)) + geom_line() + geom_point() + geom_text(aes(label = trmonth, vjust=-1)) + geom_line(color ="red", data = a11lm_train_predict, aes(y = mean_amount_pred))
```

```{r}
a11lm_train_predict <- data.frame(mean_amount_pred = predict(a_i1_l1_lm_train, a_i1_l1), trdate = a_i1_l1$trdate)

ggplot(a_i1_l1, aes( x = trdate, y = mean_amount)) + geom_line() + geom_point() + geom_text(aes(label = trmonth, vjust=-1)) + geom_line(color ="red", data = a11lm_train_predict, aes(y = mean_amount_pred))
```

```{r}
a_i1_l1_lm_bad <- lm(mean_amount ~ gdp + CPI * shols + trmonth, data = a_i1_l1, na.action = na.omit)
summary(a_i1_l1_lm_bad)
plot(a_i1_l1_lm_bad)

a11lm_predict <- data.frame(mean_amount_pred = predict(a_i1_l1_lm_bad, a_i1_l1), trdate = a_i1_l1$trdate)

a_i1_l1_plot + geom_line(color ="red", data = a11lm_predict, aes(y = mean_amount_pred))
```

```{r}
x = model.matrix(~ ., a_i1_l1[, c(-2, -9)])
y = a_i1_l1$mean_amount

cv.fit_ridge = cv.glmnet(x, y, family = 'multinomial')
```

