---
title: "AT1_A_MODELLING_1"
author: "Joshua McCarthy"
date: "16/04/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, eval = FALSE}
install.packages("lubridate")
```


```{r}
library(lubridate)
library(tidyverse)
```



```{r Prep}
#import
transactions <- read_csv("transactions.csv")

# convert dates to dates
transactions$date <- as.Date(transactions$date, format = "%d/%m/%Y")

#change variable name to avoid conflicts
tname <- names(transactions)
tname[1] <- "trdate"
names(transactions) <- tname

# convert neccessary variables to factors
transactions$customer_id <- as.factor(transactions$customer_id) 
transactions$industry <- as.factor(transactions$industry)
transactions$location <- as.factor(transactions$location)
```

# 2. Basic Model Fitting

## a. Creating the model

### i. Create an aggregated data set using the fields date, industry and location, with a mean of monthly_amount.

```{r}
aggregated_transactions <- group_by(transactions, trdate, location, industry) %>%
  summarise(mean_amount = mean(monthly_amount))
```

### ii. Create a line plot of the variable monthly_amount for industry = 1 and location = 1.  Note the seasonality by month in this time series. 

```{r}
aggregated_transactions$trmonth <- month(aggregated_transactions$trdate, label = TRUE, abbr = TRUE)
```


```{r}
a_i1_l1 <- aggregated_transactions %>%
  filter(location == 1, industry == 1)

a_i1_l1_plot <- ggplot(a_i1_l1 , aes( x = trdate, y = mean_amount)) + geom_line() + geom_point() + geom_text(aes(label = trmonth, vjust=-1))

a_i1_l1_plot
```

### iii. For industry= 1 and location = 1, train a linear regression model with monthly_amount as the target. 

Intial Model test
```{r}
a_i1_l1_lm_bad <- lm(mean_amount ~ trdate, a_i1_l1)
summary(a_i1_l1_lm_bad)
plot(a_i1_l1_lm_bad)

a11lm_predict <- data.frame(mean_amount_pred = predict(a_i1_l1_lm_bad, a_i1_l1), trdate = a_i1_l1$trdate)

a_i1_l1_plot + geom_line(color ="red", data = a11lm_predict, aes(y = mean_amount_pred))
```

#### i. Note 1 :Remember that time is very important in this model, so be sure to include variable(s) for the time sequence.

```{r}
# EDA
a_i1_l1_filt <- a_i1_l1 %>%
  data.frame() %>%
  select(mean_amount)
ts_a_i1_l1 <- ts(a_i1_l1_filt, start = decimal_date(as.Date("2013-01-01")), frequency = 12)
plot(ts_a_i1_l1)
components_a_i1_l1 <- decompose(ts_a_i1_l1)
plot(components_a_i1_l1)
```

```{r}
a_i1_l1 <- a_i1_l1 %>%
  data.frame() %>%
  select(-location, -industry)
```


```{r}
a_i1_l1_lm_bad <- lm(mean_amount ~., a_i1_l1)
summary(a_i1_l1_lm_bad)
plot(a_i1_l1_lm_bad)

a11lm_predict <- data.frame(mean_amount_pred = predict(a_i1_l1_lm_bad, a_i1_l1), trdate = a_i1_l1$trdate)

a_i1_l1_plot + geom_line(color ="red", data = a11lm_predict, aes(y = mean_amount_pred))
```


```{r}
a_i1_l1$quaters <- "Q4"
  a_i1_l1[a_i1_l1$trmonth == "Jan" | a_i1_l1$trmonth == "Feb" | a_i1_l1$trmonth == "Mar" , "quaters"] <- "Q1"
  a_i1_l1[a_i1_l1$trmonth == "Apr" | a_i1_l1$trmonth == "May" | a_i1_l1$trmonth == "Jun" , "quaters"] <- "Q2"
  a_i1_l1[a_i1_l1$trmonth == "Jul" | a_i1_l1$trmonth == "Aug" | a_i1_l1$trmonth == "Sep" , "quaters"] <- "Q3"
a_i1_l1$quaters <- as.factor(a_i1_l1$quaters)  
```


```{r}
a_i1_l1_lm_bad <- lm(mean_amount ~ trdate + quaters, a_i1_l1)
summary(a_i1_l1_lm_bad)
plot(a_i1_l1_lm_bad)

a11lm_predict <- data.frame(mean_amount_pred = predict(a_i1_l1_lm_bad, a_i1_l1), trdate = a_i1_l1$trdate)

a_i1_l1_plot + geom_line(color ="red", data = a11lm_predict, aes(y = mean_amount_pred))
```
```{r}
a_i1_l1_lm_bad <- lm(mean_amount ~ trdate + quaters + trmonth, a_i1_l1)
summary(a_i1_l1_lm_bad)
plot(a_i1_l1_lm_bad)

a11lm_predict <- data.frame(mean_amount_pred = predict(a_i1_l1_lm_bad, a_i1_l1), trdate = a_i1_l1$trdate)

a_i1_l1_plot + geom_line(color ="red", data = a11lm_predict, aes(y = mean_amount_pred))
```

```{r}
# inflation 
# https://www.statista.com/statistics/271845/inflation-rate-in-australia/
inflation <- tibble(year = c(2013, 2014, 2015, 2016), inflperc = c(2.45, 2.51, 1.51, 1.25))

# gdp growth
# https://data.worldbank.org/indicator/NY.GDP.MKTP.KD.ZG?end=2016&locations=AU&start=2013
gdpgrowth <- tibble(year = c(2013, 2014, 2015, 2016), growthperc = c(2.585,2.533,2.193,2.771))
gdpgrowth <- gdpgrowth %>%
  slice(rep(1:n(), each = 12))

gdpgrowth$growthperc <- gdpgrowth$growthperc/12/100
gdpgrowth <- gdpgrowth[-48,]
gdpgrowth$growthperccumul <- gdpgrowth$growthperc
gdpgrowth <- data.table(gdpgrowth)

for (inter in 2:nrow(gdpgrowth)) {
  gdpgrowth[inter,3] <- gdpgrowth[[inter,2]] + gdpgrowth[[inter-1,3]]
}
```


```{r}
# CPI
# https://www.rba.gov.au/inflation/measures-cpi.html
df <- data.frame(cpi = c(102.0 ,102.4 ,102.8 ,104.0 ,104.8 ,105.4 ,105.9 ,106.4 ,106.6 ,106.8 ,107.5 ,108.0 ,108.4 , 108.2 ,108.6 ,109.4 ,110.0 ), d = (1:17))

plot(df$d, df$cpi)
```


```{r}
# vs USD

# vs EURO

# vs YUAN
```

```{r}
a_i1_l1$gdp <- gdpgrowth$growthperccumul
plot(a_i1_l1$trdate, a_i1_l1$gdp)
```


```{r}
a_i1_l1_lm_bad <- lm(mean_amount ~ trmonth + gdp, a_i1_l1)
summary(a_i1_l1_lm_bad)
plot(a_i1_l1_lm_bad)

a11lm_predict <- data.frame(mean_amount_pred = predict(a_i1_l1_lm_bad, a_i1_l1), trdate = a_i1_l1$trdate)

a_i1_l1_plot + geom_line(color ="red", data = a11lm_predict, aes(y = mean_amount_pred))
```


* Holidays
* School Holidays
* Seasons
* Inflation
* Population growth
* Economic Growth
* Dollar

##### (Hint: on your plot you may see local trends like seasonality. Consider how you could craft a variable to capture this?).

##### You may also see a global upwards or downwards slope, could you craft a variable to capture this? Therefore there are two simple variables you could create to capture time.

##### Could you craft more complex ones, perhaps with polynomials to capture local or global trends? Experiment and see! 

#### ii. Note 2: Carefully think about how you split your test and train sets. (Hint: Random is not appropriate!)

Splitting will be;
* Training set - 3 earliest years (01/01/2013 - 31/12/2015)
* Testing set - most recent year(01/01/2016 - end)

```{r}
a_i1_l1_trainset <- a_i1_l1 %>%
  filter(trdate <= "15-12-31")

a_i1_l1_testset <- a_i1_l1 %>%
  filter(trdate > "15-12-31")
```

