---
title: "junk"
output: html_notebook
---

```{r}

customers <- tibble(cid = unique(transactions$customer_id), duplicates = NA)
doubles <- NA


# for (variable in customers$cid) {
if (
transactions2 %>%
  filter(customer_id == "b6d0d3c61122345690a19bb8af632db4") %>%
  nrow() 

==

transactions2 %>%
  filter(customer_id == "b6d0d3c61122345690a19bb8af632db4") %>%
  filter(!duplicated(date)) %>%
  nrow()
) { 
 rbind(doubles, "FALSE")
} else { 
  rbind(doubles, "TRUE")
  }
```

```{r}
industries_and_plots <- data.frame(stringsAsFactors = FALSE)
for (indust in 1:10){
  dfname <- paste("transactions_industry", indust, sep = "_")
  plotname <- paste("plot_trans_industry", indust, sep = "_")
  dfind <- transactions %>%
    filter(industry == indust)
  plotind <- ggplot(dfind, aes(x=date , y= monthly_amount)) + geom_point(alpha = 0.25)
  assign(dfname, dfindust)
  assign(plotname, plotind)
  ind_row <- cbind(dfname, plotname)
  industries_and_plots <- rbind(industries_and_plots, ind_row)
}
```

```{r}
industries_and_plots <- data.frame(stringsAsFactors = FALSE)
plotlist_industry <- list()
for (indust in 1:10){
  dfname <- paste("transactions_industry", indust, sep = "_")
  plotname <- paste("plot_trans_industry", indust, sep = "_")
  dfind <- transactions %>%
    filter(industry == indust)
  plotind <- ggplot(dfind, aes(x=date , y= monthly_amount)) + geom_point(alpha = 0.25)
  assign(dfname, dfindust)
  assign(plotname, plotind)
  plotlist_industry[[indust]] <- plotname
}
```

```{r}
plotlist_industry <- lapply(list, function)
  plotind <- ggplot(dfind, aes(x=date , y= monthly_amount)) + geom_point(alpha = 0.25)
  assign(plotname, plotind)
plotlist_industry[[indust]] <- plotname

industries_and_plots$plotname <- as.character(industries_and_plots$plotname)
plotlist_industry <- as.list(industries_and_plots$plotname)

# print(as.list(industries_and_plots$plotname))
# multiplot(plot_trans_industry_1,plot_trans_industry_2,plot_trans_industry_3, cols = 2)



testlist <- as.list(plot_trans_industry_1,plot_trans_industry_2,plot_trans_industry_3)
testlist[[]] <- plot_trans_industry_4

```

```{r}
```{r}
# modified code from CANVAS Announcement
# industry_location is the aggregated data frame
plotlist_loc_ind <- data.frame()
industries = unique(transactions$industry)
locations = unique(transactions$location)
x = 1
for (ind in industries) {
for (loc in locations) {
  temp_name <- paste("plot", x, sep = "_")
# create a subset of the data
temp = transactions[transactions$industry == ind &
transactions$location == loc, ]
assign(temp_name, temp)
row <- cbind(temp_name)
plotlist_loc_ind <- rbind(plotlist_loc_ind, row)
x <- x+1
}
}

plotlist_loc_ind <- as.list(1:100)

plotlist_loc_ind <- lapply(plotlist_loc_ind, function(int){
  ggplot(eval(as.symbol(paste("plot_", int, sep=""))), aes(x = date, y = monthly_amount)) + geom_point(alpha = 0.25) + theme_void() + labs (title = paste("plot", int, sep = " "))
  })

multiplot(plotlist = plotlist_loc_ind, cols = 5)
```

```{r}
# aggregated_transactions <- transactions %>%
 # select(trdate, industry, location, monthly_amount)

# aggregated_transactions <- aggregate(aggregated_transactions, list(aggregated_transactions$trdate), mean)
```

```{r}
agg_ind1_loc1$xmasbreak <- "no"
agg_ind1_loc1[agg_ind1_loc1$month == "Dec" | agg_ind1_loc1$month == "Jan", "xmasbreak"] <- "yes"

agg_ind1_loc1_lm <- lm(mean_amount ~ trdate + xmasbreak, agg_ind1_loc1)
summary(agg_ind1_loc1_lm)
plot(agg_ind1_loc1_lm)

a11lm_predict <- data.frame(mean_amount_pred = predict(agg_ind1_loc1_lm, agg_ind1_loc1), trdate = agg_ind1_loc1$trdate)

agg_ind1_loc1_plot + geom_line(color ="red", data = a11lm_predict, aes(y = mean_amount_pred))
```

```{r Time Series Stuff}
ts_a_i1_l1_adj <- ts_a_i1_l1 - components_a_i1_l1$seasonal
ts_a_i1_l1_adj_stat <- diff(ts_a_i1_l1_adj, differences = 1)
plot(ts_a_i1_l1_adj_stat)
acf(ts_a_i1_l1_adj_stat, lag.max = 12)
```

